---
title: "R Project 2"
author: "Sitong Chen"
date: "9/21/2021"
output: "html_document"
---

```{r, message=FALSE}
library(readr)
library(tidyverse)
library(ggplot2)
library(data.table)
library(reshape2)
library(table1)
```

```{r}
movies <- read.csv("./Data/movies.csv", stringsAsFactors=FALSE) #You may need to download the csv file from Data directory
ratings <- read.csv("./Data/ratings.csv")
str(movies)
```

```{r}
summary(movies)
head(movies)
```

```{r}
summary(ratings)
head(ratings)
```
#### Combine ratings and movies together by movieId and create some graphs
```{r}
movie_rating <- merge(ratings,movies,by="movieId")
```

#### Clean the date
```{r}
library(lubridate)
tibble('Initial Date' = date(as_datetime(min(movie_rating$timestamp),origin="1970-01-01")),
       'Final Date' = date(as_datetime(max(movie_rating$timestamp),origin="1970-01-01"))) %>%
  mutate(Period = duration(max(movie_rating$timestamp)-min(movie_rating$timestamp)))
```

```{r}
library(ggthemes)
library(scales)
movie_rating %>% mutate(year = year(as_datetime(timestamp, origin="1970-01-01"))) %>%
  ggplot(aes(x=year)) +
    geom_histogram(color = "white") + 
    ggtitle("Rating Distribution Per Year") +
    xlab("Year") +
    ylab("Number of Ratings") +
    scale_y_continuous(labels = comma) + 
    theme_economist()
```

#### Rating Distribution
```{r}
summary(movie_rating$rating)
# Count the number for each ratings
rating_dist <- movie_rating %>% group_by(rating) %>% summarize(n=n())
g <- ggplot(rating_dist, aes(x=rating, y=n))+geom_col()+labs(title="Ratings Distribution", y="Count",x="Ratings")
g
```

#### Investigate the users to see how many users rate over 1000 movies
##### 15 userid rate over 1000 movies
```{r}
movie_rating %>% group_by(userId) %>%
  summarise(n=n()) %>% 
  arrange(n) %>%
  filter(n>=1000)
```

#### Distribution of users with their rating count
```{r}
movie_rating %>% group_by(userId)%>%
  summarise(n=n()) %>%
  ggplot(aes(n)) + 
  geom_histogram(color = "white") + 
  scale_x_log10()+
  ggtitle("Distribution of Users")+
  xlab("Number of Ratings") +
  ylab("Number of Users")+
  scale_y_continuous(labels=comma)+
  theme_economist()


```



##### I see there is a machine learning approach to study these datasets and create a movie recommendation system. I'm currently taking a machine learning course and I would like to try it out!
#### Create a genre table with all the genres listed in movies dataset
##### I need to convert the genres present in the movies dataframe into a more usable format by the users.
##### To create a one-hot encoding to create a matrix that composes of corresponding genres for each of the films. 
```{r}
movies_genre <- as.data.frame(movies$genres, stringAsFactors = FALSE)
movies_genre2 <- as.data.frame(tstrsplit(movies_genre[,1],'[|]',type.convert = TRUE),stringAsFactors = FALSE)
colnames(movies_genre2) <- c(1:10)
genre_list <- c("Action", "Adventure", "Animation", "Children", 
                "Comedy", "Crime","Documentary", "Drama", "Fantasy",
                "Film-Noir", "Horror", "Musical", "Mystery","Romance",
                "Sci-Fi", "Thriller", "War", "Western")
genre_mat1 <- matrix(0,10330,18)
genre_mat1[1,]<-genre_list
colnames(genre_mat1) <- genre_list

for (index in 1:nrow(movies_genre2)){
  for (col in 1:ncol(movies_genre2)){
    genre_col = which(genre_mat1[1,]==movies_genre2[index,col])
    genre_mat1[index+1, genre_col] <- 1
  }
}

genre_mat2 <- as.data.frame(genre_mat1[-1,],stringAsFactors = FALSE) #remove the first row of genre list

for (col in 1:ncol(genre_mat2)){
  genre_mat2[,col] <- as.integer(genre_mat2[,col])
}

str(genre_mat2)
```


#### Create a search matrix that can perform an easy search of the films by specifying the genre present in our list. 
```{r}
searchmatrix <- cbind(movies[,1:2],genre_mat2)
head(searchmatrix)
```


#### To be continued.....
